---
title: "Lab 01 - Hello R"
author: "Lily Botha"
date: "1/15/2026"
output: 
  html_document: 
    fig_height: 3
    fig_width: 6
    theme: lumen
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(datasauRus)
data("datasaurus_dozen")
```

## Exercises

### Exercise 1
1846 rows, 3 columns (dataset, x, and y).

Make a frequency table of dataset variable. 
```{r}
data("datasaurus_dozen")
datasaurus_dozen %>%
  count(dataset)%>%
  print()
```

### Exercise 2

Filter for observations where dataset=dino, store as new frame named "dino_data". Visualize/plot data. 
```{r plot-dino, fig.height=3, fig.width=6}
dino_data <- datasaurus_dozen %>%
  filter(dataset == "dino")

ggplot(data = dino_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

Calculate correlation coefficient between x and y. 
```{r cor-dino}
dino_data %>%
  summarize(r = cor(x, y))
```

### Exercise 3

Filter for observations where dataset=star, store as new frame named "star_data". Visualize/plot data. 
```{r plot-star}
star_data <- datasaurus_dozen %>%
  filter(dataset == "star")

ggplot(data = star_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

Calculate correlation coefficient between x and y
```{r cor-star}
star_data %>%
  summarize(r = cor(x, y))
```
The dino and star plot have the same correlation coefficient. 

### Exercise 4

Filter for observations where dataset=circle, store as new frame named "circle_data". Visualize/plot data.
```{r plot-circle}
circle_data <- datasaurus_dozen %>%
  filter(dataset == "circle")

ggplot(data = circle_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

Calculate correlation coefficient between x and y. 
```{r cor-circle}
circle_data %>%
  summarize(r = cor(x, y))
```
The dino, star, and circle plots all have very similar correlation coefficients. 

### Exercise 5

Facet by the dataset variable, then place plots in 3 column grid. Then generate all correlation coefficients. 

```{r}
ggplot(datasaurus_dozen, aes(x = x, y = y, color = dataset)) +
  geom_point() +
  facet_wrap(~dataset, ncol = 3) +
  theme(legend.position = "none")

datasaurus_dozen %>%
  group_by(dataset) %>%
  summarize(r = cor(x, y)) %>%
  print(13)
```



