---
title: "Lab 01 - Hello R"
author: "Lily Botha"
date: "1/15/2026"
output: 
  html_document: 
    fig_height: 4
    fig_width: 9
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(datasauRus)
data("datasaurus_dozen")
```

## Exercises

### Exercise 1
1846 rows, 3 columns (dataset, x, and y).

```{r}
# make a frequency table of dataset variable
data("datasaurus_dozen")
datasaurus_dozen %>%
  count(dataset)%>%
  print()
```

### Exercise 2

```{r plot-dino, fig.height=3, fig.width=6}
# filter for observations where dataset = dino, store as new frame named "dino_data"
dino_data <- datasaurus_dozen %>%
  filter(dataset == "dino")

# visualize/plot data
ggplot(data = dino_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

```{r cor-dino}
# calculate correlation coefficient between x and y
dino_data %>%
  summarize(r = cor(x, y))
```

### Exercise 3

```{r plot-star}
# filter for observations where dataset = star, store as new frame named "star_data"
dino_data <- datasaurus_dozen %>%
  filter(dataset == "star")

# visualize/plot data
ggplot(data = star_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

```{r cor-star}
# calculate correlation coefficient between x and y
dino_data %>%
  summarize(r = cor(x, y))
```
The dino and star plot have the same correlation coefficient. 

### Exercise 4

```{r plot-circle}
# filter for observations where dataset = circle, store as new frame named "circle_data"
circle_data <- datasaurus_dozen %>%
  filter(dataset == "circle")

# visualize/plot data
ggplot(data = circle_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

```{r cor-circle}
# calculate correlation coefficient between x and y
circle_data %>%
  summarize(r = cor(x, y))
```
The dino, star, and circle plots all have very similar correlation coefficients. 

### Exercise 5

This is your final exercise! Add code and narrative as needed. Be creative and show off what you’ve learned so far.
Add code and narrative as needed.

To add additional R chunks, you can either type the backticks, curly braces, and the letter `r`, or use the "Insert Chunk" button above (the green C+ icon). Have fun!


